---
title: "Social Inequality"
subtitle: "Trends, International Differences and Societal Impacts"
author: "Aymen Feki, Aziz Gheni, Fadi Bouachour, Moez Daghfous"
institute: "Dr. Hoffmann"
date: "2025-12-18"
format:
  revealjs:
    css: customstyle.css
    embed-resources: true
    slide-number: c
    logo: "Results/LMU_Logo.svg"
editor: visual
---

```{r setup, include=FALSE, warning=FALSE}
source("env_setup.R")
source("Program/main.R")
```

# Agenda

::: incremental
1.  Overview and Terminology

2.  Data Analysis

    Dynamics and International Differences in Social Inequality

    Societal Correlates of Social Inequality

3.  Summary and Discussion
:::

# Research Questions {.custom-smaller}

Understanding how social inequality evolves over time and how it interacts with key societal outcomes is essential for assessing economic fairness, social cohesion, and institutional stability.\
Against this background, this project addresses the following questions:

::: incremental
1.  How has social inequality evolved over time within individual countries and at the global level?

2.  To what extent do countries differ in their levels of social inequality?

3.  How is social inequality related to other societal factors?
:::

# Overview {.custom-smaller2}

-   **Data type:** Longitudinal country–year panel data (2000–2023)

-   **Source:** Our World in Data

-   **Coverage:** Global country coverage with an unbalanced panel structure due to heterogeneous data availability.

-   **Datasets:**

    -   Income Inequality
    -   Gender Inequality Index
    -   Economist Intelligence Unit Democracy Index
    -   Homicide Rate
    -   Universal Health Coverage Service Coverage Index
    -   World Population

# Terminology {.custom-smaller}

::: incremental
-   **Gini Index :** Income Inequality *\[0 = equal, 1 = unequal\]*
-   **Gender Inequality Index (GII) :** Gender Inequality (Health, Empowerment and Labor) *\[0 = equal, 1 = unequal\]*
-   **Economist Intelligence Unit Democracy Index (EIU-DI) :** Measures democratic quality *\[0–10, higher = more democratic\]*
-   **Universal Health Coverage Service Coverage Index (UHC-SCI) :** Evaluates essential health services *\[0-100, higher = better coverage\]*
:::

# Agenda

[1. Overview and Terminology]{style="color:lightgray;"}\

2.  Data Analysis

    Dynamics and International Differences in Social Inequality\

    [Societal Correlates of Social Inequality]{style="color:lightgray;"}\

[3. Summary and Discussion]{style="color:lightgray;"}\

# Income Inequality

## Gini Index (After Tax)

```{r gini_density, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gini.density
```

## Continental Changes Over Time

```{r gini_continents_plot, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gini.continents.plot
```

## International Changes Over Time

```{r gini_countries_plot, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gini.countries.plot
```

## Cross-Country Comparison

```{r gini_countries_boxplot, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gini.countries.boxplot
```

# Gender Inequality

## Gender Inequality Index

```{r gii_density, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gii.density
```

## International Changes Over Time

```{r gii_countries_plot, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gii.countries.plot
```

## Cross-Country Comparison

```{r gii_countries_boxplot, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gii.countries.boxplot
```

# Agenda

[1. Overview and Terminology]{style="color:lightgray;"}\

2.  Data Analysis

    [Dynamics and International Differences in Social Inequality]{style="color:lightgray;"}\

    Societal Correlates of Social Inequality

[3. Summary and Discussion]{style="color:lightgray;"}\

## Democracy

```{r scatter_gini_democracy, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.gini.democracy
```

## Democracy

```{r scatter_highlight_gini_democracy, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.highlight.gini.democracy
```

## Homicide Rate

```{r scatter_gini_homicide, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.gini.homicide.rate
```

## Homicide Rate

```{r scatter_highlight_gini_homicide, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.highlight.gini.homicide.rate
```

## Healthcare Access

```{r scatter_gini_healthcare, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.gini.healthcare.access
```

## Healthcare Access

```{r scatter_highlight_gini_healthcare, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.highlight.gini.healthcare.access
```

# Agenda

[1. Overview and Terminology]{style="color:lightgray;"}

[2. Data Analysis]{style="color:lightgray;"}

3.  Summary and Discussion

# Summary {.custom-smaller2}

::: incremental
-   Countries exhibit markedly different trajectories and rates of change in inequality over time.

-   The overall level of inequality differs substantially across countries.

-   Inequality indicators display distinct distributions across countries and over time.

-   Weak negative monotonic relationship between income inequality and the Democracy Index.

-   Strong positive monotonic relationship between income inequality and homicide rate.

-   Moderate negative monotonic relationship between income inequality and healthcare access.
:::

# Discussion {.custom-smaller2}

::: incremental
-   **Data quality and reporting issues:**

    -   Political incentives may affect the reporting of democracy, crime, and inequality.
    -   Authoritarian regimes may underreport negative outcomes.
    -   Survey-based inequality measures can miss top incomes.

-   **Index simplifications:** Composite indices (Gini, Democracy, Healthcare) provide simplified views and cannot capture all dimensions of social inequality.

-   **Sensitivity to influential countries:** Correlations may be driven by a small number of extreme cases.

-   **Uneven data coverage:** Missing values and unbalanced time series reduce cross-country comparability.
:::
