---
title: "Social Inequality"
subtitle: "Trends, International Differences and Societal Impacts"
author: "Aymen Feki, Aziz Gheni, Fadi Bouachour, Moez Daghfous"
institute: "Dr. Hoffmann"
date: "2025-12-18"
format:
  revealjs:
    css: customstyle.css
    embed-resources: true
    slide-number: c
    logo: "Results/LMU_Logo.svg"
editor: visual
---

```{r setup, include=FALSE, warning=FALSE}
source("env_setup.R")
source("Program/functions.R")
source("Program/load_data.R")
source("Program/Main.R")
```

# Agenda

::: incremental
1.  Overview and Terminology

2.  Data Analysis

    Dynamic and International Differences in Social Inequality

    Societal Correlates of Social Inequality

3.  Summary and Discussion
:::

# Overview

::: incremental
-   Trends analyzed from 2000 to 2023.

-   Development of Inequality within and across countries and continents analyzed using Indices.

-   Relationship between Income Inequality and societal outcomes.
:::

# Terminology {.custom-smaller}

::: incremental
-   **Gini Index :** Income Inequality *\[0 = equal, 1 = unequal\]*
-   **Gender Inequality Index (GII) :** Gender Inequality (Health, Empowerment and Labor) *\[0 = equal, 1 = unequal\]*
-   **Democracy Index :** Measures democracy *\[0–10, higher = more democratic\]*
-   **Universal Health Coverage Service Coverage Index (UHCSCI) :** Evaluates essential health services *\[0-100, higher = better coverage\]*
:::

# Agenda

[1. Overview and Terminology]{style="color:lightgray;"}\

2.  Data Analysis

    Dynamic and International Differences in Social Inequality\

    [Societal Correlates of Social Inequality]{style="color:lightgray;"}\

[3. Summary and Discussion]{style="color:lightgray;"}\

# Income Inequality

## Gini Index

```{r gini_density, echo=FALSE, fig.align='center', fig.dpi=100, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gini.density
```

## Continental Changes Over Time

```{r gini_continents_plot, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gini.continents.plot
```

## International Changes Over Time

```{r gini_countries_plot, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gini.countries.plot
```

## Cross-Country Comparison

```{r gini_countries_boxplot, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gini.countries.boxplot
```

# Gender Inequality

## Gender Inequlity Index

```{r gii_density, echo=FALSE, fig.align='center', fig.dpi=100, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gii.density
```

## International Changes Over Time

```{r gii_countries_plot, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gii.countries.plot
```

## Cross-Country Comparison

```{r gii_countries_boxplot, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
gii.countries.boxplot
```

# Agenda

[1. Overview and Terminology]{style="color:lightgray;"}\

2.  Data Analysis

    [Dynamic and International Differences in Social Inequality]{style="color:lightgray;"}\

    Societal Correlates of Social Inequality

[3. Summary and Discussion]{style="color:lightgray;"}\

## Democracy

```{r scatter_gini_democracy, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.gini.democracy
```

## Democracy

```{r scatter_highlight_gini_democracy, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.highlight.gini.democracy
```

## Homicide Rate

```{r scatter_gini_homicide, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.gini.homicide
```

## Homicide Rate

```{r scatter_highlight_gini_homicide, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.highlight.gini.homicide
```

## Healthcare Access

```{r scatter_gini_healthcare, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.gini.healthcare
```

## Healthcare Access

```{r scatter_highlight_gini_healthcare, echo=FALSE, fig.align='center', fig.dpi=300, fig.height=4, fig.width=7, message=FALSE, warning=FALSE}
scatter.highlight.gini.healthcare
```

# Agenda

[1. Overview and Terminology]{style="color:lightgray;"}

[2. Data Analysis]{style="color:lightgray;"}

3.  Summary and Discussion

# Summary

::: incremental
-   Social Inequality evolves very differently across countries over time.

-   Moderate correlation between the Gini coefficient and the democracy index.

-   Noticeable association between the Gini coefficient and the homicide rates.

-   Non-negligible negative relationship between Income Inequality and access to healthcare.
:::

# Discussion

::: incremental
-   Possible data Inaccuracies due to political factors and the sensitivity of the subject

-   The Indices used are simplifications and do not fully capture the complexity of social Inequality.

-   The observed correlations may be driven by a few influential countries, which could distort the overall relationships.

-   Missing and Insufficient data for several countries
:::
